---
title: "Mountain Race"
description: |
  Animation of my Strava efforts on one of my local climbs
author:
  - name: Julian During
date: "`r Sys.Date()`"
output: html_document
repository_url: https://github.com/duju211/mountain_race
creative_commons: CC BY
base_url: https://www.datannery.com/posts/mountain-race/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("libraries.R")
```

# Idea

Every cyclist has a particular important climb.
It might not be a big deal to anyone else, but any climb can be important!

My favorite local climb goes by the name of 'Lochen'.
It's located outside of my local
hometown Balingen in the southwest of Germany. It's about 4.4 kilometers long
with an average gradient of 6.9%.

This doesn't sound like a hard climb. It might not even register as a
regular big climb for most cyclists. But for me it's one of the most iconic
climbs.

In the following post, I will let different versions of me race against each
other on my favorite local climb!

In order to reproduce the analysis, perform the following steps:

* Clone the [repository](https://github.com/duju211/mountain_race)
* Install the packages listed in the `libraries.R` file
* Run the target pipeline by executing `targets::tar_make()` command


# Data

The data originates from my personal Strava data. 
If you have a Strava account and want to query your data like I do here, you
can have a look at one of my
[previous posts](https://www.datannery.com/posts/strava-data/).

The data are a bunch of arrow files, that you can via dpylr syntax
thanks to the DuckDB package.

Deselect `heartrate` measurements and restrict the spatial data to a bounding
box. Add information about the type and the start date of each
activity.

```{r lochen, file="https://raw.githubusercontent.com/duju211/pin_strava/master/R/poi.R"}
```

```{r, echo=FALSE}
tar_read(df_lochen_raw)
```

Further preprocess the raw data. Keep only rows, where I was moving and turn
the start date from datetime to date. Adjust
the `time` column so that every activity starts at time 0.


```{r vis_lochen, file="R/lochen.R"}
```

```{r, echo=FALSE}
tar_read(df_lochen)
```


# Visualisation

Make a first static ggplot visualisation. Keep the plot rather minimal.
Use `ggplot2::theme_void` as a general theme:

```{r, file="R/vis_lochen.R"}
```

```{r gg_lochen, echo=FALSE, preview=TRUE}
tar_read(gg_lochen_raw)
```

As you can see there are lot of paths on one road. These are my
bike rides on the 'Lochen' pass.

Some paths don't seem to match. These are activities of another
type in the same region as my bike rides.
These activities don't use the main road and stand out in the
plot.

To further explore the data, make a first animated visualisation with
the `gganimate` package:

```{r, file="R/vis_anim_lochen.R"}
```

```{r, echo=FALSE}
knitr::include_graphics(tar_read(gif_anim_lochen))
```

In this animated version of the plot, you can see that there are further
problems in the data. Not all bike rides start at the bottom of the climb.
You can guess which activities start at the top of the climb, by looking 
at the general speed of the animation :). Determine these activities:

```{r, file="R/wrong_direction.R"}
```

Filter the activities for bike rides.
Exclude activities that start at the top of the climb. Repeat the above
animated plot:

```{r, file="R/lochen_ride.R"}
```

```{r, echo=FALSE}
knitr::include_graphics(tar_read(gif_anim_lochen_ride))
```

Now it looks much cleaner and the rides are more comparable to one another.

For the final version of the animation, add small points to point out my
position at the time. Color these positions by the `year` of the activity
Reduce the
speed of the animation a little bit, to display smaller differences in the
rides.

```{r, file="R/vis_anim_lochen_final.R"}
```

```{r, echo=FALSE}
knitr::include_graphics(tar_read(gif_anim_lochen_ride_final))
```

I very much like how the plot turned out. I hope I can add many more layers
to this animation in the future!